---
title: "Migrating database from MySQL to Neo4j"
author: "Yun Mai"
date: "April 30, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NoSQL migration

Assignment:

For this assignment, you should take information from a relational database and migrate it to a NoSQL database of your own choosing. 

For the relational database, you might use the flights database, the tb database, the "data skills" database your team created for Project 3, or another database of your own choosing or creation.

For the NoSQL database, you may use MongoDB (which we introduced in week 7), Neo4j, or another NoSQL database of your choosing.

Your migration process needs to be reproducible.  R code is encouraged, but not required.  You should also briefly describe the advantages and disadvantages of storing the data in a relational database vs. your NoSQL database.

Overview:

I expanded the movie rating database created on week2 homework in MySQL. subsequently, I migrated the movie_rating database to Neo4j.

```{r,eval=F}
devtools::install_github("nicolewhite/Rneo4j") 
```

```{r}
library(RMySQL)
library(RNeo4j)
```

## Get the Data from MySQL
```{r, echo= F}
mysql_pw <- "5201"
```

```{r}
# connect to MySQL local database
movie_con <- dbConnect(MySQL(), user="root", password=mysql_pw, dbname = "movie_rating", host = "localhost")
```

```{r}
#running query
query.omdb <- dbSendQuery(movie_con, "SELECT * FROM omdb")
#retrieving the omdb data from MySQL
omdb.data <- dbFetch(query.omdb, n=-1)

#send query to MySQL and retrieve the movie data from MySQL
query.movie <- dbSendQuery(movie_con, "SELECT * FROM movie")
movie.data <- dbFetch(query.movie, n=-1)

#send query to MySQL and retrieve the friends data from MySQL
query.friends <- dbSendQuery(movie_con, "SELECT * FROM friends")
friends.data <- dbFetch(query.friends, n=-1)

#send query to MySQL and retrieve the rating data from MySQL
query.rating <- dbSendQuery(movie_con, "SELECT * FROM rating")
rating.data <- dbFetch(query.rating, n=-1)

#send query to MySQL and retrieve the actor data from MySQL
query.actor <- dbSendQuery(movie_con, "SELECT * FROM actor")
actor.data <- dbFetch(query.actor, n=-1)

#send query to MySQL and retrieve the director data from MySQL
query.director <- dbSendQuery(movie_con, "SELECT * FROM director")
director.data <- dbFetch(query.director, n=-1)

#send query to MySQL and retrieve the writer data from MySQL
query.writer <- dbSendQuery(movie_con, "SELECT * FROM writer")
writer.data <- dbFetch(query.writer, n=-1)

#send query to MySQL and retrieve the genre data from MySQL
query.genre <- dbSendQuery(movie_con, "SELECT * FROM genre")
genre.data <- dbFetch(query.genre, n=-1)

#send query to MySQL and retrieve the country data from MySQL
query.country <- dbSendQuery(movie_con, "SELECT * FROM country")
country.data <- dbFetch(query.country, n=-1)
```

**Export data as .csv files**
```{r}
write.csv(omdb.data, "omdb.csv")
write.csv(movie.data, "movie.csv")
write.csv(friends.data, "friends.csv")
write.csv(rating.data, "rating.csv")
write.csv(actor.data, "actor.csv")
write.csv(director.data, "director.csv")
write.csv(writer.data, "writer.csv")
write.csv(genre.data, "genre.csv")
write.csv(country.data, "country.csv")

```

**Disconnect from MySQL**
```{r}
dbDisconnect(movie_con)
```


## Load the data to Neo4j

```{r,echo=F}
neo4j.pw <- "data607"
```

```{r}
#connect to Neo4j
graph = startGraph("http://localhost:7474/db/data", username = "neo4j", password = neo4j.pw)

#add the necessary uniqueness constraints
addConstraint(graph, "Movie", "title")
addConstraint(graph, "Person", "name")
addConstraint(graph, "Genre", "genre")
addConstraint(graph, "Country", "country")
 



```

